---
title: "prep_DSNY_2"
output: html_document
---
```{r}
library(dplyr)
library(readr)
library(janitor)
library(tidyverse)
```

 # DSNY Monthly Tonnage Data - https://data.cityofnewyork.us/City-Government/DSNY-Monthly-Tonnage-Data/ebb7-mvp5
 
```{r}
tonnage_raw <- read_csv("DSNY_Monthly_Tonnage_Data.csv")
```

```{r}
tonnage_clean <- clean_names(tonnage_raw)
```
 # arranging data by year 
```{r}
tonnage_sorted <- arrange(tonnage_clean, month)
```
 # splitting the column to separate year from month 
```{r}
adj_tonnage <- separate(tonnage_sorted, month, into = c("years", "months"),sep = "/")
```
 # filtering to keep data from 2015 - 2020 
```{r}
recent_tonnage <- filter(adj_tonnage, years > 2015)
recent_tonnage <- filter(recent_tonnage, years < 2021)
```
 # group and summarize to see volume of trash by borough
```{r}
boro_tonnage <- group_by(recent_tonnage, borough_id)
boro_tonnage <- summarize(boro_tonnage, total_tonnage = sum(refusetonscollected))
boro_tonnage <- ungroup(boro_tonnage)
boro_tonnage
```

 # have to filter by borough and then community district (CD) to compare CD refuse produced 

 # filtering out manhattan CD total refuse 

```{r}
mh_tonnage <- filter(recent_tonnage, borough_id ==1)
mh_dist_tonnage <- group_by(mh_tonnage, communitydistrict)
mh_dist_tonnage <- summarize(mh_dist_tonnage, total_tonnage = sum(refusetonscollected))
mh_dist_tonnage <- ungroup(mh_dist_tonnage)
mh_dist_tonnage
```

 # filtering out brooklyn CD total refuse 

```{r}
bk_tonnage <- filter(recent_tonnage, borough_id ==3)
bk_dist_tonnage <- group_by(bk_tonnage, communitydistrict)
bk_dist_tonnage <- summarize(bk_dist_tonnage, total_tonnage = sum(refusetonscollected))
bk_dist_tonnage <- ungroup(bk_dist_tonnage)
bk_dist_tonnage
```

 # filtering out queens CD total refuse 
 
```{r}
 qn_tonnage <- filter(recent_tonnage, borough_id ==4)
 qn_dist_tonnage <- group_by(qn_tonnage, communitydistrict)
 qn_dist_tonnage <- summarize(qn_dist_tonnage, total_tonnage = sum(refusetonscollected))
 qn_dist_tonnage <- ungroup(qn_dist_tonnage)
 qn_dist_tonnage
```

 # filtering out bronx CD total refuse 

```{r}
 bx_tonnage <- filter(recent_tonnage, borough_id ==2)
 bx_dist_tonnage <- group_by(bx_tonnage, communitydistrict)
  bx_dist_tonnage <- summarize(bx_dist_tonnage, total_tonnage = sum(refusetonscollected))
 bx_dist_tonnage <- ungroup(bx_dist_tonnage)
 bx_dist_tonnage
```

 # filtering out for staten island CD total refuse 

```{r}
st_tonnage <- filter(recent_tonnage, borough_id ==5)
st_dist_tonnage <- group_by(st_tonnage, communitydistrict)
st_dist_tonnage <- summarize(st_dist_tonnage, total_tonnage = sum(refusetonscollected))
st_dist_tonnage <- ungroup(st_dist_tonnage)
st_dist_tonnage
```

 # sorting data to evaluate total paper recyclables collected by borough 

```{r}
boro_papers <- group_by(recent_tonnage, borough_id)
boro_papers <- summarize(boro_papers, total_papers = sum(papertonscollected))
boro_papers <- ungroup(boro_papers)
boro_papers
```

 # filtering out paper recycling by manhattan community districts

```{r}
mh_papers <- filter(recent_tonnage, borough_id ==1)
mh_dist_papers <- group_by(mh_papers, communitydistrict)
mh_dist_papers <- summarize(mh_dist_papers, total_papers = sum(papertonscollected))
mh_dist_papers <- ungroup(mh_dist_papers)
mh_dist_papers
```

  # filtering out paper recycling by brooklyn community districts

```{r}
bk_papers <- filter(recent_tonnage, borough_id ==3)
bk_dist_papers <- group_by(bk_papers, communitydistrict)
bk_dist_papers <- summarize(bk_dist_papers, total_tonnage = sum(papertonscollected))
bk_dist_papers <- ungroup(bk_dist_papers)
bk_dist_papers
```

 # filtering out paper recycling by queens community districts 

```{r}
 qn_papers <- filter(recent_tonnage, borough_id ==4)
 qn_dist_papers <- group_by(qn_papers, communitydistrict)
 qn_dist_papers <- summarize(qn_dist_papers, total_tonnage = sum(papertonscollected))
 qn_dist_papers <- ungroup(qn_dist_papers)
 qn_dist_papers
```

 # filtering out bronx recycling by community district 

```{r}
bx_papers <- filter(recent_tonnage, borough_id ==2)
bx_dist_papers <- group_by(bx_papers, communitydistrict)
bx_dist_papers <- summarize(bx_dist_papers, total_tonnage = sum(papertonscollected))
bx_dist_papers <- ungroup(bx_dist_papers)
bx_dist_papers
```

 # filtering out staten island paper recycling by CD

```{r}
st_papers <- filter(recent_tonnage, borough_id ==5)
st_dist_papers <- group_by(st_papers, communitydistrict)
st_dist_papers <- summarize(st_dist_papers, total_tonnage = sum(papertonscollected))
st_dist_papers <- ungroup(st_dist_papers)
st_dist_papers
```

 # creating new column summing up total waste, total_waste

```{r}
recent_tonnage2 <- mutate(recent_tonnage, 
                          total_waste = papertonscollected + refusetonscollected + mgptonscollected)
```

```{r}
write_csv(boro_tonnage, "C:\\Users\\Michelle\\Desktop\\boro_tonnage.csv")
```
```{r}
write_csv(recent_tonnage2, "C:\\Users\\Michelle\\Desktop\\full_dsny_clean.csv")
```

 # breaking out total waste and total paper recycling by borough 

```{r}
boro_waste <- group_by(recent_tonnage2, borough_id)
boro_waste <- summarize(boro_waste, total_waste_boro = sum(total_waste))
boro_waste <- ungroup(boro_waste)
boro_waste

boro_papers <-group_by(recent_tonnage2, borough_id)
boro_papers <- summarize(boro_papers, total_papers = sum(papertonscollected))
boro_papers <- ungroup(boro_papers)
boro_papers
```

```{r}
write_csv(boro_waste, "C:\\Users\\Michelle\\Desktop\\total_waste_by_boro.csv")
write_csv(boro_papers, "C:\\Users\\Michelle\\Desktop\\total_paper_rec_by_boro.csv")
```

```{r}
write_csv(mh_dist_papers, "C:\\Users\\Michelle\\Desktop\\mh_total_paper_rec_by_dist.csv")
write_csv(mh_dist_tonnage, "C:\\Users\\Michelle\\Desktop\\mh_total_tonnage_by_dist.csv")

write_csv(bx_dist_papers, "C:\\Users\\Michelle\\Desktop\\bx_total_paper_rec_by_dist.csv")
write_csv(bx_dist_tonnage, "C:\\Users\\Michelle\\Desktop\\bx_total_tonnage_by_dist.csv")

write_csv(bk_dist_papers, "C:\\Users\\Michelle\\Desktop\\bk_total_paper_rec_by_dist.csv")
write_csv(bk_dist_tonnage, "C:\\Users\\Michelle\\Desktop\\bk_total_tonnage_by_dist.csv")

write_csv(qn_dist_papers, "C:\\Users\\Michelle\\Desktop\\qn_total_paper_rec_by_dist.csv")
write_csv(qn_dist_tonnage, "C:\\Users\\Michelle\\Desktop\\qn_total_tonnage_by_dist.csv")

write_csv(st_dist_papers, "C:\\Users\\Michelle\\Desktop\\st_total_paper_rec_by_dist.csv")
write_csv(st_dist_tonnage, "C:\\Users\\Michelle\\Desktop\\st_total_tonnage_by_dist.csv")
```

 # creating new column, rec_total, to represent total recyclables collected 
 
```{r}
recent_tonnage3 <- mutate(recent_tonnage2, 
                          rec_total = papertonscollected +  mgptonscollected)
```

```{r}
write_csv(recent_tonnage3, "C:\\Users\\Michelle\\Desktop\\total_data_updated.csv")
```

 # breaking out total trash (x_totals) and total recyclables (x_recs)by CD

```{r}
st_totals <- filter(recent_tonnage3, borough_id ==5)
st_totals <- group_by(st_totals, communitydistrict)
st_totals <- summarize(st_totals, total_trash = sum(total_waste))
st_totals <- ungroup(st_totals)
st_totals
```

```{r}
st_recs <- filter(recent_tonnage3, borough_id ==5)
st_recs <- group_by(st_recs, communitydistrict)
st_recs <- summarize(st_recs, total_rec = sum(rec_total))
st_recs <- ungroup(st_recs)
st_recs
```

```{r}
bx_totals <- filter(recent_tonnage3, borough_id ==2)
bx_totals <- group_by(bx_totals, communitydistrict)
bx_totals <- summarize(bx_totals, total_trash = sum(total_waste))
bx_totals <- ungroup(bx_totals)
bx_totals
```

```{r}
bx_recs <- filter(recent_tonnage3, borough_id ==2)
bx_recs <- group_by(bx_recs, communitydistrict)
bx_recs <- summarize(bx_recs, total_rec = sum(rec_total))
bx_recs <- ungroup(bx_recs)
bx_recs
```

```{r}
bk_totals_dist <- filter(recent_tonnage3, borough_id ==3)
bk_totals_dist <- group_by(bk_totals_dist, communitydistrict)
bk_totals_dist <- summarize(bk_totals_dist, total_trash = sum(total_waste))
bk_totals_dist <- ungroup(bk_totals_dist)
bk_totals_dist
```
```{r}
bk_recs_dist <- filter(recent_tonnage3, borough_id ==3)
bk_recs_dist <- group_by(bk_recs_dist, communitydistrict)
bk_recs_dist <- summarize(bk_recs_dist, total_rec = sum(rec_total))
bk_recs_dist <- ungroup(bk_recs_dist)
bk_recs_dist
```
```{r}
mh_totals_dist <- filter(recent_tonnage3, borough_id ==1)
mh_totals_dist <- group_by(mh_totals_dist, communitydistrict)
mh_totals_dist <- summarize(mh_totals_dist, total_trash = sum(total_waste))
mh_totals_dist <- ungroup(mh_totals_dist)
mh_totals_dist
```
```{r}
mh_recs_dist <- filter(recent_tonnage3, borough_id ==1)
mh_recs_dist <- group_by(mh_recs_dist, communitydistrict)
mh_recs_dist <- summarize(mh_recs_dist, total_rec = sum(rec_total))
mh_recs_dist <- ungroup(mh_recs_dist)
mh_recs_dist
```
```{r}
qn_totals_dist <- filter(recent_tonnage3, borough_id ==4)
qn_totals_dist <- group_by(qn_totals_dist, communitydistrict)
qn_totals_dist <- summarize(qn_totals_dist, total_trash = sum(total_waste))
qn_totals_dist <- ungroup(qn_totals_dist)
qn_totals_dist
```
```{r}
qn_recs_dist <- filter(recent_tonnage3, borough_id ==4)
qn_recs_dist <- group_by(qn_recs_dist, communitydistrict)
qn_recs_dist <- summarize(qn_recs_dist, total_rec = sum(rec_total))
qn_recs_dist <- ungroup(qn_recs_dist)
qn_recs_dist
```



